\texttt{thornado}-transport: IMEX schemes for two-moment neutrino transport respecting Fermi-Dirac statistics

### 2019-02-01 Documents/M1/Astronum_2018/sections/Conclusion.tex
Work on Astronum paper.

We have developed IMEX schemes suitable for a two-moment model of neutrino transport that obey Fermi-Dirac statistics.
The scheme employs algebraic closure based on Fermi-Dirac statistics, high-order discontinuous Galerkin methods for spatial discretization, and convex-invariant time integration to maintain realizability of the moments.  
Since the realizable domain is convex and its convexity can be inherited by a convex combination, a scheme having convex combinations as its stages can preserve the realizable domain.
This encouraged us to construct realizability-preserving time integrators, realizability-preserving IMEX schemes, and eventually a method with a realizability-preserving IMEX time integrator and high-order DG method.

In the applications that motivate this work, the neutrino distribution function can vary from 0 to 1.  
Hence, we have considered algebraic closures based on Fermi-Dirac statistics for both low and high occupancy.  
Among the seven algebraic closures we considered -- Kershaw~\cite{kershaw_1976}, Wilson~\cite{wilson_1975,leblancWilson_1970}, Levermore~\cite{levermore_1984}, Minerbo~\cite{minerbo_1978}, Janka 1~\cite{janka_1991}, Janka 2~\cite{janka_1992}, and Cernohorsky \& Bludman~\cite{cernohorskyBludman_1994} -- only the Cernohorsky \& Bludman closure obeys Fermi-Dirac statistics for all occupancies.  
As a result, we employed the Cernohorsky \& Bludman closure for the neutrino stationary state test in Section~\ref{se: Neutrino Stationary State Test}.
(We also ran the code with other algebraic closures.
And the program crashed after a few time steps because of the nonphysical neutrino density/flux obtained and no correction step was taken for $f>1$.
We could make the code run by adding correction step, but the quality of the results may be compromised.
And it's difficult to determine whether the results are in fact compromised by a particular correction step and, if so, in what way and to what degree.)

Two PD-ARS schemes are proposed.
The one with SSPRK2 has second-order accuracy while the other with SSPRK3 has third-order accuracy, and both have the strong-stability preserving property in the streaming limit.  
Their accuracy was demonstrated on problems with known smooth solutions in streaming, absorption, and scattering-dominated regimes.
The neutrino transport test with emission, absorption, and isoenergetic scattering through a stationary background, was designed to test the convex-invariance of our PD-ARS schemes. 
The neutrino stationary state test shows that a method combining an algebraic closure based on Fermi-Dirac statistics and convex-invariant time integration can be a candidate for constructing a CCSN simulation code without ``unphysical correction steps" that may break conservation laws.

In this work, we adopted Cartesian coordinates, a linear collision term, and a fixed non-relativistic material background.
More realistic problems of scientific interest, such as with energy-exchanging scattering and relativistic effects, are left for future research.

>>>>

We have developed IMEX schemes suitable for a two-moment model of neutrino transport that obey Fermi-Dirac statistics.
The scheme employs algebraic closure based on Fermi-Dirac statistics, high-order discontinuous Galerkin methods for spatial discretization, and convex-invariant time integration to maintain realizability of the moments.  
Since the realizable domain is convex and its convexity can be inherited by a convex combination, a scheme having convex combinations as its stages can preserve the realizable domain.
This encouraged us to construct realizability-preserving time integrators, realizability-preserving IMEX schemes, and a method with a realizability-preserving IMEX time integrator and high-order DG method.

In the applications that motivate this work, the neutrino distribution function can vary from 0 to 1.  
Hence, we have considered algebraic closures based on Fermi-Dirac statistics for both low and high occupancy.  
Among the seven algebraic closures we considered -- Kershaw~\cite{kershaw_1976}, Wilson~\cite{wilson_1975,leblancWilson_1970}, Levermore~\cite{levermore_1984}, Minerbo~\cite{minerbo_1978}, Janka 1~\cite{janka_1991}, Janka 2~\cite{janka_1992}, and Cernohorsky \& Bludman~\cite{cernohorskyBludman_1994} -- only the Cernohorsky \& Bludman closure obeys Fermi-Dirac statistics for all occupancies.  
As a result, we employed the Cernohorsky \& Bludman closure for the neutrino stationary state test in Section~\ref{se: Neutrino Stationary State Test}.
We also ran our code with Minerbo closure, IMEX PC2, IMEX SSP2332 and IMEXRKCB2 schemes.
And the results show that only PD-ARS schemes have stability.
In addition, closures have impact on the simulation result.
Even though RKCB2 with Minerbo closure luckily survived in our test, the result it gives is compromised.
In what way and what degree the results are in fact compromised either by the closure or by a particular correction step for unrealizable moments are difficult to determine and left for further study.

Two PD-ARS schemes are proposed.
The one with SSPRK2 has second-order accuracy while the other with SSPRK3 has third-order accuracy, and both have the strong-stability preserving property in the streaming limit.  
Their accuracy was demonstrated on problems with known smooth solutions in streaming, absorption, and scattering-dominated regimes.
The neutrino transport test with emission, absorption, and isoenergetic scattering through a stationary background, was designed to test the convex-invariance of our PD-ARS schemes. 
The neutrino stationary state test shows that a method combining an algebraic closure based on Fermi-Dirac statistics and convex-invariant time integration is promising for robust CCSN simulation.

In this work, we adopted Cartesian coordinates, a linear collision term, and a fixed material background.
More realistic problems of scientific interest, such as with energy-exchanging scattering and relativistic effects, are left for future research.

### 2019-02-01 Documents/M1/Astronum_2018/sections/TwoMoments.tex
Work on Astronum paper.

However, recently reported CCSN simulations using two-moment neutrino transport with algebraic closures have employed Eddington factors that can violate the bounds in Eq.~\eqref{eq:eddingtonFactorBounds}.  
As examples, we consider the Eddington factors discussed in~\cite{murchikova_etal_2017}, where the suitability of several algebraic closures for two-moment neutrino transport was evaluated.  
In Fig.~\ref{fig:EddingtonFactorsWithDifferentClosure}, we plot the Eddington factor versus the flux factor for two occupancies; $\cJ = 0.1$ (low occupancy) and $\cJ = 0.9$ (high occupancy).  
Of the algebraic closures plotted, few satisfy the bounds on the Eddington factor in Eq.~\eqref{eq:eddingtonFactorBounds}.  
Kershaw~\cite{kershaw_1976}, Wilson~\cite{wilson_1975,leblancWilson_1970}, Levermore~\cite{levermore_1984}, Minerbo~\cite{minerbo_1978}, and Janka~2~\cite{janka_1992} closures may work fine when the occupancy is low.  
When the occupancy is high, the Eddington factor due to these closures exceeds the upper bound for Fermi-Dirac statistics.  
The Eddington factor of Janka~1~\cite{janka_1991} may violate both the upper and lower bound on the Eddington factor.  
Only the closure due to Cernohorsky~\&~Bludman~\cite{cernohorskyBludman_1994} satisfies both the upper and lower bounds.  
This is not surprising as this is the only closure based on Fermi-Dirac statistics.  
Although the Levermore and Minerbo closures do not satisfy the bounds in Eq.~\eqref{eq:eddingtonFactorBounds}, they are widely used in simulations of neutrino transport in CCSNe and compact binary mergers; e.g., O'Connor \& Couch~\cite{oConnorCouch_2018}, Pan et al.~\cite{pan_etal_2018}, Glas et al.~\cite{glas_etal_2018}, Just et al.~\cite{just_etal_2018}, and Foucart et al.~\cite{foucart_etal_2015} use the Minerbo closure, while Vartanyan et al.~\cite{vartanyan_etal_2018}, Cabezon et al.~\cite{cabezon_etal_2018}, Kuroda et al.~\cite{kuroda_etal_2016}, and Fujibayashi et al.~\cite{fujibayashi_etal_2017} use the Levermore closure.  
When employing these closures in conditions of high occupancy, a numerical scheme may evolve the moments outside the realizable domain of Fermi-Dirac statistics given in Eq.~\eqref{eq:MomentsBounds}.
If this were to happen, the update step may give $\cJ>1$.
Considering the fact that the completed collision term contains blocking factors, i.e.~$(1-\cJ)\times$ something positive, $\cJ > 1$ would change blocking factors' sign and it would be difficult to predict the impact of the subsequent induced errors on the simulation outcome.
\rc{Besides, even with the simplified collision term, Eq~\ref{eq:collisionTermMoments}, $\cJ>1$ could result... the reason of crash}
If treatments are developed to map the unrealizable moments into the realizable domain, they should be developed to conserve lepton number, energy, and momentum.

>>>>

However, recently reported CCSN simulations using two-moment neutrino transport with algebraic closures have employed Eddington factors that can violate the bounds in Eq.~\eqref{eq:eddingtonFactorBounds}.  
As examples, we consider the Eddington factors discussed in~\cite{murchikova_etal_2017}, where the suitability of several algebraic closures for two-moment neutrino transport was evaluated.  
In Fig.~\ref{fig:EddingtonFactorsWithDifferentClosure}, we plot the Eddington factor versus the flux factor for two occupancies; $\cJ = 0.1$ (low occupancy) and $\cJ = 0.9$ (high occupancy).  
Of the algebraic closures plotted, few satisfy the bounds on the Eddington factor in Eq.~\eqref{eq:eddingtonFactorBounds}.  
Kershaw~\cite{kershaw_1976}, Wilson~\cite{wilson_1975,leblancWilson_1970}, Levermore~\cite{levermore_1984}, Minerbo~\cite{minerbo_1978}, and Janka~2~\cite{janka_1992} closures may work fine when the occupancy is low.  
When the occupancy is high, the Eddington factor due to these closures exceeds the upper bound for Fermi-Dirac statistics.  
The Eddington factor of Janka~1~\cite{janka_1991} may violate both the upper and lower bound on the Eddington factor.  
Only the closure due to Cernohorsky~\&~Bludman~\cite{cernohorskyBludman_1994} satisfies both the upper and lower bounds.  
This is not surprising as this is the only closure based on Fermi-Dirac statistics.  
Although the Levermore and Minerbo closures do not satisfy the bounds in Eq.~\eqref{eq:eddingtonFactorBounds}, they are widely used in simulations of neutrino transport in CCSNe and compact binary mergers; e.g., O'Connor \& Couch~\cite{oConnorCouch_2018}, Pan et al.~\cite{pan_etal_2018}, Glas et al.~\cite{glas_etal_2018}, Just et al.~\cite{just_etal_2018}, and Foucart et al.~\cite{foucart_etal_2015} use the Minerbo closure, while Vartanyan et al.~\cite{vartanyan_etal_2018}, Cabezon et al.~\cite{cabezon_etal_2018}, Kuroda et al.~\cite{kuroda_etal_2016}, and Fujibayashi et al.~\cite{fujibayashi_etal_2017} use the Levermore closure.  
When employing these closures in conditions of high occupancy, a numerical scheme may evolve the moments outside the realizable domain of Fermi-Dirac statistics given in Eq.~\eqref{eq:MomentsBounds}.
If this were to happen, the update step may give $\cJ>1$.
Considering the fact that the completed collision term contains blocking factors, i.e.~$(1-\cJ)\times$ something positive, $\cJ > 1$ would change blocking factors' sign and it would be difficult to predict the impact of the subsequent induced errors on the simulation outcome.
Besides, even with the simplified collision term Eq~\ref{eq:collisionTermMoments}, $\cJ>1$ could result information lose (such as, lose the neutrino distribution for high-energy groups), closure failure, and other unpredictable consequences.
If treatments are developed to map the unrealizable moments into the realizable domain, they should be developed to conserve lepton number, energy, and momentum.
